Discrete Ordinates

### 2017-06-08 docprep/chapimpl.tex
Updated thesis text

This chapter covers the details of the implementation of the different components of the DOCTORS code base.

The code is available from a GitHub Git repository. At the time of this writing, the repository is located at \texttt{https://github.com/eNorris/thesis} and is only available subject to special request and license agreement. An executable version is available.

@@

Data is stored as \texttt{std::vector<>} objects.

@@

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/indx_ex}
  \end{center}
  \caption{The indexing scheme.}
\label{fig:indx_ex}
\end{figure}

@@

\section{Cross Section Parsing}\label{sec:xsparse}
AMPX cross sections are stored in a binary format. The binary file is split into a series of records. Each record contains a string of binary bytes between a four byte header and four byte footer. The header and footer are identical and, when interpreted as a signed four byte integer, give the size in words of the record. Each word is a 32-bit section (4~bytes) of binary data. An example record is shown in Fig.~\ref{fig:ampxbytes1}.

The data is stored in Big Endian format which must be converted to Little Endian for most Intel and AMD processors.  The endianess rearrangement is shown in Fig.~\ref{fig:ampxbytes2}. All records are one of 12 unique types. Each type of record is formatted and interpreted differently. For example, Type 1 contains general information about the data file such as the number of nuclides, number of energy groups, and a brief description. Type 2 records contains a list of floating-point numbers representing the energy group boundaries. Figure~\ref{fig:ampxbytes1} shows a Type 1 record. In the case of the header bytes given in Fig.~\ref{fig:ampxbytes1} and~\ref{fig:ampxbytes2}, the bytes correspond to the integer 440 which is the length in bytes of the record (not including the header or footer).

Each word of data is sequentially read, reordered, and converted to an apprpriately typed variable. Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/ampxbytes1}
  \end{center}
  \caption{An example record parsed. The header and footer are identical and are the number of 32-bit words in the record.}
\label{fig:ampxbytes1}
\end{figure}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=2.75in]{figs/ampxbytes2}
  \end{center}
  \caption{The byte reordering from Big Endian to Little Endian. Individual bits withing a byte are not rearranged, but the ordering of the four bytes withing a 32-bit word is reversed.}
\label{fig:ampxbytes2}
\end{figure}

@@

DOCTORS reads AMPX fromatted binary data files. Figure~\ref{fig:ampx} shows the data format of the overall file.

@@

The 2D data is stored in a special format optimized for scatter matrix data.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/ampx1}
        \caption{The AMPX file format}
        \label{fig:ampx1}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/ampx2}
        \caption{The structure of each individual record.}
        \label{fig:ampx2}
    \end{subfigure}
    \caption{The format of the AMPX file. (a) The file contains a header, a list of directory records, energy group information and a list of nuclide entries. (b) Each nuclide entry contains multiple records; the two of concern in this work are the last two containing gamma data.}\label{fig:ampx}
\end{figure}

Figure~\ref{fig:sigmacomp} shows the microscopic cross section data pulled from

@@

Once the data is parsed, weighted combinations of elemental data is used to create material cross sections.

MT AND MF NUMBERS

WATER EXAMPLE.

@@

\section{Qt5 Framework}\label{sec:qt}
The Qt5 framework was used for implementation of the user interface. Qt5 enables asynchronous calls through its ssignal/slot mechanism. A signal can be emitted which will execute all connected slots. Signals and slots can be connected manually by the user except for ones automatically generated by Qt. Qt prvides a user interface for building user interfaces. Components such a sbuttons, drop boxes, radio buttons, etc. are offered to the user. Qt makes extensive use of polymorphism. All user components inherit from the base \texttt{QWidget} class which inherits from \texttt{QObject}. Any class that utilizes the signal/sloot mechanism must extend the \texttt{QOjbect} class.

@@

\begin{listing}
\begin{minted}[frame=lines,linenos]{python}
import numpy as np
def myfcn(hello,world):
   return hello*world
\end{minted}
\caption{Blah blah blah.}\label{lst:sync1}
\end{listing}

@@

@@

The GUI is cool. Lots of pictures of it.

>>>>

This chapter covers the details of the implementation of the different components of the DOCTORS code base. The code is available from a GitHub Git repository. At the time of this writing, the repository is located at \texttt{https://github.com/eNorris/thesis} and is only available subject to special request and license agreement. An executable version is available.

@@

Data is stored in large, multi-dimensional arrays. For example, the CT voxel phantom can be intuitively stored as a 3D array of floating precision values which can be easily indexed. However, DOCTORS is written such that it is extensible to GPUs. GPUs are not optimized for data stored in a multi-dimensional fashion, but rather for data stored as a single 1-dimensional array. To emulate the 3-dimensional storage arrangement of the data, indexing arithmetic is used.

On the CPU, data is stored as C++ \texttt{std::vector<T>} objects where the \texttt{T} template parameter can be either \texttt{float} or \texttt{double} depending on the precision (32 bits or 64 bits respectively) needed. The storage space available to a \texttt{std::vector<T>} object is effectively unlimited, bounded only by the host CPU's physical RAM. However, other effects such as memory fragmentation can limit the practical size of a single, continuous \texttt{std::vector<T>} object significantly. To emulate indexing of a $N_x \times N_y$ matrix, a single \texttt{std::vector<T>} is created and resized to $N_x \times N_y$ elements. The global index, $i$, in the flattened array is then
\begin{equation}
i = i_x N_y + i_y.
\end{equation}
This pattern extends to multi-dimensional matrices. For example, a $N_x \times N_y$ matrix would be globally indexed as
\begin{equation}
i = i_x N_y N_z + i_y N_z + i_z.
\end{equation}

@@

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/indx_ex}
  \end{center}
  \caption{Conversion from 3D indexing to 1D "flattened" indexing.}
\label{fig:indx_ex}
\end{figure}

@@

\section{Cross Section Parsing}\label{sec:xsparse}
AMPX cross sections are stored in a binary format. Figure~\ref{fig:ampx} shows the data format of the overall file. The binary file is split into a sequence of records. The header section of the file has overall information about the file including the number of nuclides stored, their energy group structure, and a comment string describing file. Following the header section is a list of directory records. Each directory has information about a specific nuclide including the cross section reaction types available, temperatures for which the cross sections were evaluated, the number of Legendre expansion coefficients stored in the scatter matrix, and other data. Following the directories, more records follow that describe the energy structure used in the file. There is one such record for each particle type. The libraries used in the current version of DOCTORS have both neutron and photon cross section data stored.

The file header section, directories, and energy group boundaries describe the structural information necessary to read the following nuclide records. Nuclides are listed in the same order their directories are found after the header setion. Each nuclide contains numerous records.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/ampx1}
        \caption{}
        \label{fig:ampx1}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/ampx2}
        \caption{}
        \label{fig:ampx2}
    \end{subfigure}
    \caption{The format of the AMPX file. (a) The file contains a header, a list of directory records, energy group information and a list of nuclide entries. (b) Each nuclide entry contains multiple records; the two of concern in this work are the last two containing gamma data.}\label{fig:ampx}
\end{figure}

Each record contains a string of data bytes between a four byte header and four byte footer. The header and footer are identical and, when interpreted as a signed four byte integer, give the size in bytes of the record. An example record is shown in Fig.~\ref{fig:ampxbytes1}.

The data is stored in Big Endian format which must be converted to Little Endian for most Intel and AMD processors.  The endianess rearrangement is shown in Fig.~\ref{fig:ampxbytes2}. Each record is one of 12 unique types; each type of record is formatted and interpreted differently. For example, Type 1 contains general information about the data file such as the number of nuclides, number of energy groups, and a brief description. Type 2 records contains a list of floating-point numbers representing the energy group boundaries. Figure~\ref{fig:ampxbytes1} shows a Type 1 record. In the case of the header bytes given in Fig.~\ref{fig:ampxbytes1} and~\ref{fig:ampxbytes2}, the bytes correspond to the integer 440 which is the length in bytes of the record (not including the header or footer).

Each byte of data is sequentially read, reordered, and converted to an apprpriately typed variable. Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/ampxbytes1}
  \end{center}
  \caption{An example record parsed. The header and footer are identical and equal to the number of bytes in the record.}
\label{fig:ampxbytes1}
\end{figure}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=2.75in]{figs/ampxbytes2}
  \end{center}
  \caption{The byte reordering from Big Endian to Little Endian. Individual bits within a byte are not rearranged, but the ordering of the four bytes withing a 32-bit word is reversed.}
\label{fig:ampxbytes2}
\end{figure}

@@

@@

The 2D data is stored in a special format optimized for scatter matrix data. The format concists of a sequence of "magic numbers" interpersed within a list of data points. A magic number is a nine digit number IIIJJJKKK. The first three digits, III, are the group number of the highest energy group to scatter to the energy of interest. The next three digits, JJJ, are the group number of the lowest energy to scatter. The final three digits, KKK, are the group number of the sink energy to which particles are scattering. Given a magic number \texttt{magic}, the three values can be computed in Listing~\ref{lst:magic}. Note that the group numbers are indexed from 1 to $G$ instead of 0 to $G-1$ as required by DOCTORS. This is corrected by subtracting 1 while doing the indexing arithmetic.

\begin{listing}
\begin{minted}[frame=lines,linenos]{c++}
# Given magic
KKK = magic % 1000
JJJ = (magic % 1000000 - KKK) /1000
III = (magic - JJJ - KKK) / 1000000

src = JJJ
while src >= III
	data = READ_NEXT_BINARY_FLOAT()
	xs[(src - 1)*G + KKK - 1] = data
	src = src - 1
\end{minted}
\caption{Computation of the source and sink groups from the magic number and the subsequent data parsing.}\label{lst:magic}
\end{listing}

Figure~\ref{fig:sigmacomp} shows the microscopic cross section data pulled from the 200-neutron/47-gamma group data file currently used by DOCTORS for hydrogen and oxygen. The data is compared to reference data pulled directly from the ENDF/B-VII.1 photoatomic (MT=501) data library~\citep{ref:cullend} accessible through the Sigma database~\citep{ref:sigma}.

@@

Once the data is parsed, weighted combinations of elemental data is used to create material cross sections. Cross section data for photons always uses the naturally ocurring since photo-atomic reactions are not sensitive to the nuclear differences between isotopes. The compositions of N materials are assumed to be given as weight fractions, $w_i$ subject to
\begin{equation}
sum_{i = 1}^N w_i = 1
\end{equation}
and is converted to an atom fraction, $a_i$:
\begin{equation}
a_i = \frac{w_i}{M m_i}
\end{equation}
where
\begin{equation}
M = \sum_{i = 1}^N \frac{w_i}{m_i}
\end{equation}
and $m_i$ is the element's atomic weight.

In the solution employed by DOCTORS, only the total and scattering cross sections are required. The AMPX data files however, support arbitrary reaction types and have up to a dozen or more reactions for photo-atomic reactions alone for elemental datasets. The reaction types are identified by their MT designation. The total, inelastic, and elastic scatter cross sections are MT 501, 502, and 504 respectively. The scatter cross section used by DOCTORS is the sum of the elastic and inelastic cross sections.

The cross section of compound materials are computed as a weighted summation of their components. For example, the cross section of water for a particular reaction would be
\begin{equation}
\sigma_{H_2 O} = \frac{2 \sigma_H + \sigma_O}{3}.
\end{equation}

@@

\section{Qt5 Framework}\label{sec:qt}
The Qt5 framework was used for implementation of the user interface. Qt5 enables asynchronous calls through its signal/slot mechanism. Signals and slots are special functions that have additional processing performed by the meta-object-compiler (MOC). A signal can be emitted which will execute all connected slots. Signals and slots are connected manually by the user except special ones automatically generated by Qt5.

Qt5 prvides a user interface for building user interfaces. Components such a sbuttons, drop boxes, radio buttons, etc. are made available to the user. Through extensive use of polymorphism, Qt simplifies the addition of graphical interactive elements called widgets. All widgets inherit from the base \texttt{QWidget} class which inherits from \texttt{QObject}. Any class that utilizes the signal/slot mechanism must extend the \texttt{QOjbect} class. As an example, a button can be created in the Qt5 user interface and named \texttt{button1}. This object will be accessible as \texttt{ui->button1}. When the user clicks on the button, the signal \texttt{button1.clicked()} will automatically be emitted. The user can connect this signal to any slot with an identical number arguments of the same type. The \texttt{button1.clicked()} signal can be connected to the slot \texttt{doStuff1()} but not \texttt{doStuff2(int)} since the arguments are not compatible.

@@

\begin{listing}
\begin{minted}[frame=lines,linenos]{c++}
class MainWindow : public QWindow
{
	// Constructor
	MainWindow(QObject *parent);

	// Other parts of the class

public slot:
	void doSomethingLong();
}

void MainWindow::doSomethingLong()
{
	// Execute a long piece of code
}

MainWindow::MainWindow(QObject *parent)
{
	// Initializations

	// When a button named button1 which was created in the Qt5 creator interface is clicked, the clicked() signal is automatically emitted which executes the doSomethingLong() function
	MainWindow.connect(ui->button1, clicked(), this, doSomethingLong())
}
\end{minted}
\caption{A long function causes the user interface to block.}\label{lst:sync1}
\end{listing}

@@

\begin{listing}
\begin{minted}[frame=lines,linenos]{c++}
class MainWindow : public QWindow
{
	// Constructor
	MainWindow(QObject *parent);

	// Other parts of the class

public slot:
	void doSomethingLong();
}

void MainWindow::doSomethingLong()
{
	// Execute a long piece of code
}

MainWindow::MainWindow(QObject *parent)
{
	// Initializations

	// When a button named button1 which was created in the Qt5 creator interface is clicked, the clicked() signal is automatically emitted which executes the doSomethingLong() function
	MainWindow.connect(ui->button1, clicked(), this, doSomethingLong())
}
\end{minted}
\caption{A long function causes the user interface to block.}\label{lst:sync1}
\end{listing}

@@

INSERT PICTURES OF THE GUI

### 2017-06-08 docprep/chapintro.tex
Updated thesis text

Since its introduction in 1973 by \citet{ref:hounsfieldg} computed tomography (CT) has become pervasive in medical diagnostics as improved algorithms and techniques give doctors access to higher quality information. Increased information results in faster and more accurate diagnosis. However, the exponentially increasing usage of CT and corresponding dose patients receive has raised concerns regarding potential long term risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}. These concerns are exemplified by the push for low dose CT.

A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.

@@

No methodology currently exists to verify that the dosimetry evaluation was accurate after the patient has undergone the procedure. This work proposes patient specific a methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam. This can also help doctors estimate spatial dose distribution to the patient to ensure no specific organ received more dose than permissible.

This work implements and analyses a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient can be computed with a full transport solution inside a CT phantom. The methodology converts a CT mesh into a voxel phantom of materials and densities. The user supplies information about the beam and a discrete ordinate method computes the flux throughout the phantom. The flux is then used to compute the dose using local energy deposition to relieve the need for secondary electron transport.

@@

In addition to rapidly computing the patient dose, another goal of DOCTORS is to present the code in a user-friendly fashion. To this end, a graphical user interface (GUI) was developed. The GUI utilizes the Qt5 graphics framework for the window, buttons, and other necessary widgets. The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicated required and completed steps. The output is then plotted graphically as well as sent to an ASCII text file for more advanced postprocessing. Utilization of Qt5 also makes the code portable, it can be compiled on either Linux or Windows operating systems with no changes to the code. More details about the GUI are included in Section~\ref{sec:gui}.

BRIEF DISCUSSION OF RESULTS

>>>>

Since its introduction in 1973 by \citet{ref:hounsfieldg} computed tomography (CT) has become pervasive in medical diagnostics as improved algorithms and techniques give doctors access to higher quality information. Increased information results in faster and more accurate diagnosis. However, the exponentially increasing usage of CT and dose patients receive from it has raised concerns of potential long term risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}. These concerns are exemplified by the push for low dose CT.

INSERT MORE LOW DOSE CT STUFF
A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.

@@

No methodology currently exists to verify that the dosimetry evaluation was accurate after the patient has undergone the procedure. This work proposes a patient specific methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam. This can also help doctors estimate spatial dose distribution in the patient to ensure no specific organ received more dose than permissible.

This work implements and analyzes a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient is computed using a full transport solution inside a CT phantom. The methodology converts a CT mesh into a voxel phantom of materials and densities. The user supplies information about the beam and a discrete ordinate method computes the flux throughout the phantom. The flux is then used to compute the dose using local energy deposition to relieve the need for secondary electron transport.

@@

In addition to rapidly computing the patient dose, another goal of DOCTORS is to present the code in a user-friendly fashion. To this end, a graphical user interface (GUI) was developed. The GUI was built using the Qt5 graphics framework for the windows, buttons, and other necessary widgets. The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicated required and completed steps. The output is then plotted graphically as well as sent to an ASCII text file for more advanced postprocessing. Utilization of Qt5 also makes the code portable, it can be compiled on either Linux or Windows operating systems with no changes to the code. More details about the GUI are included in Section~\ref{sec:gui}.

INSERT BRIEF DISCUSSION OF RESULTS

### 2017-06-08 docprep/chapmeth.tex
Updated thesis text

ADD THEORY AND MATH OF ANGLE/WEIGHT GENERATION?

@@

\subsection{Anisotropy Treatment}

@@

%\begin{equation} \label{eq:col3}
%\begin{split}
%	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
%	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
%	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, %\hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', %\hat{\Omega}') dE' d\hat{\Omega}'
%\end{split}
%\end{equation}

@@

The alternative way to compute the dose is directly from the local energy deposition. The total deposition, $\mathcal{E_{T,i}}$ for voxel $i$ is
\begin{equation}\label{eq:dept}
\mathcal{E}_{T,i} = \mathcal{E}_{a,i} + \mathcal{E}_{s,i}
\end{equation} 
where $\mathcal{E_{a,i}}$ is the energy deposition from absorption and $\mathcal{E_{s,i}}$ is the energy deposition from scatter. The absorption deposition is
\begin{equation}\label{eq:depa}
\mathcal{E}_{a,i} = \sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g V_i \bar{E}^g F_1
\end{equation}
where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F_1$ is a unit conversion factor that converts MeV to Joules. The average group energy is assumed to be the average between the two bounding energies of the group. Equation~\ref{eq:depa} assumes total local energy deposition in the voxel of interest. The energy deposition from scatter is
\begin{equation}\label{eq:deps}
\mathcal{E}_{s,i} = \sum_{g=0}^{G-1} \sum_{h=g-1}^{G-1} \left[ \Sigma_{s,i}^{g,h} \varphi_i^g V_i (\bar{E}^g-\bar{E}^h) F_1 \right]
\end{equation}
making the total deposition
\begin{equation}\label{eq:dept2}
\mathcal{E}_{T,i} = V_i F_1 \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g+1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right]
\end{equation}
The mass of the voxel, $M_i$ is then computed
\begin{equation}
M_i = \rho_i V_i F_2
\end{equation}
where $\rho_i$ is the density in g/cm$^3$ and $F_2$ is a conversion factor to convert grams to kilograms. The dose (in Gy) to the voxel is then
\begin{equation}\label{eq:dosedep}
D_i = \frac{\mathcal{E}_{T,i}}{M_i} = \frac{F_1}{\rho_i F_2} \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g+1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right].
\end{equation}

>>>>

The $S_N$ quadrature ensures 90$\deg$ symmetry about any axis. In order to ensure the rotational symmetry, the $S_N$ quadrature has a single degree of freedom. The $LQ_N$ quadrature (which is the specific $S_N$ implemented in DOCTORS) uses the remaining degree of freedom to minimize error caused when integrating Legendre polynomials since that class of polynomials are used to represent anisotropy in scattering (see Sec.~\ref{sec:aniso}). For a more through discussion of quadrature generation, see \citet{ref:lewise}.

@@

\subsection{Anisotropy Treatment}\label{sec:aniso}

@@

@@

The alternative way to compute the dose is directly from the local energy deposition. The total deposition, $\mathcal{E}_{T,i}$ for voxel $i$ is
\begin{equation}\label{eq:dept}
\mathcal{E}_{T,i} = \mathcal{E}_{a,i} + \mathcal{E}_{s,i}
\end{equation} 
where $\mathcal{E}_{a,i}$ is the energy deposition from absorption and $\mathcal{E}_{s,i}$ is the energy deposition from scatter. The absorption deposition is
\begin{equation}\label{eq:depa}
\mathcal{E}_{a,i} = \sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g V_i \bar{E}^g F_1
\end{equation}
where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F_1$ is a unit conversion factor that converts MeV to Joules. The average group energy is assumed to be the average between the two bounding energies of the group. Equation~\ref{eq:depa} assumes total local energy deposition in the voxel of interest. The energy deposition from scatter is
\begin{equation}\label{eq:deps}
\mathcal{E}_{s,i} = \sum_{g=0}^{G-1} \sum_{h=g+1}^{G-1} \left[ \Sigma_{s,i}^{g,h} \varphi_i^g V_i (\bar{E}^g-\bar{E}^h) F_1 \right]
\end{equation}
making the total deposition
\begin{equation}\label{eq:dept2}
\mathcal{E}_{T,i} = V_i F_1 \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g+1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right]
\end{equation}
The mass of the voxel, $M_i$ is then computed
\begin{equation}
M_i = \rho_i V_i F_2
\end{equation}
where $\rho_i$ is the density in g/cm$^3$ and $F_2$ is a conversion factor to convert grams to kilograms. The dose (in Gy) to the voxel is then
\begin{equation}\label{eq:dosedep}
D_i = \frac{\mathcal{E}_{T,i}}{M_i} = \frac{F_1}{\rho_i F_2} \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g+1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right].
\end{equation}

### 2017-06-08 docprep/chapresults.tex
Minor changes in notation
Updates to thesis
Updated thesis text

\section{Preprocessing}
Two computational phantoms were used for verification of the DOCTORS code.

@@

The analytical point is computed as
\begin{equation}
\frac{d\varphi}{dE} = \frac{e^{-(\mu^g_a x_a + \mu^g_w x_w)}}{4\pi G (x_a + x_w)^2 (\Delta E)^g}
\end{equation}
where $\mu_a^g$ and $\mu_w^g$ are the attenuation coefficients for air and water repectively for the $g$ group, $x_a$ and $x_w$ are the pathlength through air and water respectively, and $\Delta E$ is the width of the $g$ group in MeV.

\section{3 Cone-Beam Water Benchmark}

Now there are three

@@

And now 16.

\section{1 Cone-beam Phantom Benchmark}

It's a person!

@@

It's a person!

@@

it's faster!

>>>>

\section{Preprocessing}
Two computational phantoms were used for verification of the DOCTORS code. The first phantom is a $256 \times 256 \times 64$ voxel mesh of a 35 cm water phantom. The water phantom is in an acrylic casing and the entire mesh measures $50 \times 50 \times 12.5$ cm. Each voxel is then a cube with sides of 1.95 mm. The mesh has a total of 4194304 voxels. The second phantom is of the same dimensions and resolution, but of a realistic human phantom. The mesh emulates a CT scan of a patient's midsection where the liver would be located.

@@

The analytical points are computed
\begin{equation}
\left(\frac{d\varphi}{dE}\right)^g = \frac{e^{-(\mu^g_a x_a + \mu^g_w x_w)}}{4\pi G (x_a + x_w)^2 (\Delta E)^g}
\end{equation}
where $\mu_a^g$ and $\mu_w^g$ are the attenuation coefficients for air and water repectively for the $g$ group, $x_a$ and $x_w$ are the pathlength through air and water respectively, and $\Delta E$ is the width of the $g$ group in MeV. Dividing by the number of groups maintains consistency with the computational codes which uniformly distribute particles across all groups.

In MCNP6, the effective dose was computed using a FMESH4 tally with associated DE and DF cards to weight the flux. The most accurate tally to compute the absorbed dose with is a F6 tally, however, those cannot be spatially distributed via a FMESH tally. Therefore, a F4 tally was used with a multiplier designed to emulate a F6 tally as closely as possible. The comparison is shown in Fig.~\ref{fig:mcnp_f4vsf6}.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/mcnp_f4vsf6}
  \end{center}
  \caption{Comparison of the absorbed dose computed with F4 and F6 tallies in MCNP6. All eight data points were located at the center of the phantom and the two computation methods were within 2\% of each other.}
\label{fig:mcnp_f4vsf6}
\end{figure}

The effective dose and absorbed doses are given in Fig.~\ref{fig:b2_dose}.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/b2_effdose}
        \caption{}
        \label{fig:b2_effdose}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/b2_absdose}
        \caption{}
        \label{fig:b2_absdose}
    \end{subfigure}
    \caption{(a) The effective dose computed using ICRP 116 fluence-to-dose conversion factors. (b) The absorbed dose.}\label{fig:b2_dose}
\end{figure}

@@

INSERT THE RESULTS FROM THE 16 BEAM BENCHMARK

@@

INSERT THE RESULTS FROM THE 16 BEAM PHANTOM BENCHMARK

@@

INSERT THE RESULTS FROM THE GPU ACCELERATION
